---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# PsyR

<!-- badges: start -->
<!-- badges: end -->

The goal of PsyR is to ...

## Installation

You can install the development version of PsyR from [GitHub](https://github.com/) with:

``` {r}
# install.packages("devtools")
devtools::install_github("garner-code/PsyR")
```

## Example

This is a basic example which shows you how to obtain between subject confidence 
intervals. All of the following examples will use the post-hoc method 
(see Bird, 2002, for details on the post-hoc method):

```{r b_example}

## load relevant packages
library(emmeans)
library(afex)
library(readxl)
library(PsyR)

# load some data
dat <- spacing
# load(spacing)?

# set emmeans option to multivariate
afex_options(emmeans_model = "multivariate")

# perform the statistical model
# check which model objects can be used
mod <- aov_ez("subj", "yield", dat, within = "spacing", between = "group")

# define some between group contrasts
emm_btwn <- emmeans(mod, "group")
con_b <- list(
  "12vs34" = c(0.5, 0.5, -0.5, -0.5), # groups 1 & 2 vs groups 3 & 4
  "1vs2" = c(1, -1, 0, 0), # and so on...
  "3vs4" = c(0, 0, 1, -1)
)
btwn_con <- contrast(emm_btwn, con_b)

# feed the contrast table into psyci() with the chosen method, family, and factor
# names
# extra documentation needed re: mapping of method to specific. Potentially
# add Sidak
psyci(model=mod, contrast_table = btwn_con, method="ph", family="b", 
      between_factors = list("group"))
# add attributes to the table about the method used to compute CIs
# think about adding the critical constant as an attribute
# specify df also as attributes

```


Here is an example, that carries on from the last, for obtaining within subject 
confidence intervals:

```{r w_example}

# define some within group contrasts.
# and get emm table of contrast effects
emm_win <- emmeans(mod, "spacing")
con_w <- list(
  "20vs40" = c(1, -1, 0),
  "20vs60" = c(1, 0, -1),
  "Quad" = c(0.5, -1, 0.5)
)
con_win <- contrast(emm_win, con_w)

# generate 95% CIs for the within subjects contrasts
psyci(model=mod, contrast_table = con_win, method="ph", family="w", 
      within_factors = list("spacing"))


```

And finally, here is an example of how to generate CIs for between x within 
contrasts, using the same post-hoc method:

```{r bw_example}

# get emms for each cell from the between x within design
emm_int <- emmeans(mod, c("group", "spacing"))

# the handy thing about emmeans is that you can use the already defined between 
# and within contrasts to generate your interaction contrasts. The extra 
# delightful thing is that, when used this way, emmeans will scale the contrasts
# appropriately so that you can interpret the estimated effect as the size of 
# the effect (aka it is scaled appropriately).

con_int <- contrast(emm_int, interaction=list(con_b, con_w))
# check mean difference option in Psy (go back to win app and doc)
# potentially add a scheffe function 
# add p values from gcr etc procedures to the table
# generate 95% CIs for the between x within subjects contrasts
psyci(model=mod, contrast_table = con_int, method="ph", family="bw", 
      within_factors = list("spacing"), between_factors=list("group"))

```


You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.

You can also embed plots, for example:

```{r pressure, echo = FALSE}
plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
