---
title: "Analysis-of-3-within-design"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{psyr-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette demonstrates how to use the `psyci()` function from the `PsyR` package to compute simultaneous confidence intervals (SCIs) for planned contrasts in a **within-subjects lexical decision priming experiment**.

```{r include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Loading Packages

To use Psy to calculate confidence intervals over contrast effects, we will use a workflow that uses functions from [afex](https://www.rdocumentation.org/packages/afex/versions/1.4-1) to implement ANOVA models, and [emmeans](https://rvlenth.github.io/emmeans/) to obtain estimated marginal means and the required contrast estimates. So let's load those packages, and any others that are helpful for our workflow.


```{r setup}
library(PsyR)
library(tidyverse)
library(emmeans)
library(afex)
library(ggplot2)
```

## The Data

The study investigates whether subliminal priming with a social category label (e.g., “unemployed” or “elderly”) automatically activates associated stereotypes. Participants (N = 15) complete a lexical decision task under three conditions:

- **Stereotypical** attributes
- **Atypical** attributes
- **Neutral** words

Priming is hypothesized to:
- Facilitate recognition of stereotypical words
- Inhibit recognition of atypical words relative to neutral words

The dependent variable is **ReactionTime** in tenths of a second.

We will define and test two planned contrasts:
- `"Stereo_vs_Other"`: Compares stereotypical attributes vs. the average of atypical and neutral words
- `"Atyp_vs_Neut"`: Compares atypical vs. neutral words


```{r}
data(priming)
head(priming)
```

Let’s also convert variables to factors for clarity.

```{r}
priming <- priming %>%
  mutate(
    Subject = as.factor(Subject),
    Prime = factor(Prime, levels = c("Stereotypical", "Atypical", "Neutral"))
  )
```

## Exploratory Visualization

```{r, fig.width=5, fig.height=4}
priming %>%
  group_by(Prime) %>%
  summarise(mean_RT = mean(ReactionTime)) %>%
  ggplot(aes(x = Prime, y = mean_RT, fill = Prime)) +
  geom_col() +
  theme_minimal() +
  labs(title = "Average Reaction Time by Prime Condition",
       y = "Reaction Time (tenths of seconds)")
```

---

## Statistical Analysis

We first fit a **repeated-measures ANOVA**.

```{r}
afex_options(emmeans_model = "multivariate")

mod <- aov_ez("Subject", "ReactionTime", priming, within = "Prime")
```

## Estimated Marginal Means and Planned Contrasts

Get the estimated marginal means and define **planned contrasts**:

```{r}
emm <- emmeans(mod, ~ Prime)

planned_contrasts <- list(
  "Stereo_vs_Other" = c(-2, 1, 1)/2,  # Stereotypical vs. average of Atypical & Neutral
  "Atyp_vs_Neut" = c(0, 1, -1)        # Atypical vs. Neutral
)

contrast_table <- contrast(emm, planned_contrasts)
```

## Compute Simultaneous Confidence Intervals

```{r}
scis <- psyci(
  model = mod,
  contrast_table = contrast_table,
  method = "bf",         # Bonferroni correction
  within_factors = list("Prime")
)
scis
```

## Visualize Confidence Intervals

```{r, fig.width = 7, fig.height = 4}
scis_df <- as.data.frame(scis)

scis_df$contrast <- c(
  "Stereotypical vs (Atypical + Neutral)", 
  "Atypical vs Neutral"
)

ggplot(scis_df, aes(x = contrast, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  labs(
    title = "Simultaneous Confidence Intervals for Planned Contrasts",
    y = "Contrast Estimate (RT difference)",
    x = "Planned Contrast"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

```

---

## Interpretation

Two planned contrasts were tested using simultaneous confidence intervals (SCIs) with Bonferroni correction:

1. **Stereotypical vs (Atypical + Neutral)**

   * Estimate: 3.5 (tenths of a second)
   * 95% SCI: \[2.09, 4.91]
   * p < .0001
   * **Interpretation:** Reaction times were significantly **faster** for stereotypical words compared to the average of atypical and neutral words. The confidence interval excludes 0, indicating a reliable facilitation effect due to priming.

2. Atypical vs Neutral

   * Estimate: 1.0 (tenths of a second)
   * 95% SCI: \[−1.32, 3.32]
   * p = 0.2983
   * **Interpretation:** While atypical words had numerically slower reaction times than neutral words, the confidence interval includes 0, indicating **no statistically significant difference**. This contrast does not provide strong evidence for an inhibitory priming effect on atypical attributes.


## References

Bird, K. D. (2002). *Confidence Intervals for Effect Sizes in Analysis of Variance*. Educational and Psychological Measurement, 62(2), 197–226.

---

## Learn More

* PsyR GitHub: [https://github.com/garner-code/PsyR](https://github.com/garner-code/PsyR)
* [UNSW Psychology Research Tools](https://www.unsw.edu.au/science/our-schools/psychology/our-research/research-tools/psy-statistical-program)
