---
title: "between_subjects_2x2_factorial-w-simple"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{between_subjects_2x2_factorial-w-simple}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
library(PsyR)
library(tidyverse)
library(emmeans)
library(afex)
library(forcats)
```

Following the example in chapter 4 of Bird (2004) Analysis of Variance via Confidence Intervals, of a coherent analysis for a 2 x 2 between subjects design with simple effects.

## Load the data

Lets load the data, make it longform, and define the factors 

```{r loaddata}

data(drive_sleepy)
head(drive_sleepy, 5)
```

```{r longform}

drive_sleepy <- drive_sleepy %>% 
  mutate(sleep_deprivation = factor(sleep_deprivation),
         sleep_deprivation = fct_recode(sleep_deprivation, "None" = "0", "High" = "12"),
         sleep_deprivation = fct_relevel(sleep_deprivation, "High", "None"),
         nvh = factor(nvh),
         nvh = fct_relevel(nvh, "high", "low"))

```

## Model 

Now run the anova

```{r model}

# perform the statistical model
mod <- aov_ez("subID", 
              "performance", 
              drive_sleepy, 
              between = c("sleep_deprivation", "nvh"))

```

## Contrasts

We assemble our contrasts a little bit differently for the SMR procedure,
as we only need to use it when using simple effects, and our construction of the 
null hypothesis will be common for our main and simple effects and interactions (Bird &
Hadzi-Pavlovic (2005), p. 353-354). 

So for the SMR procedure we put all our contrasts of interest into a set of lists,
and then we define our SMR parameters based on our family of models (A(b), B(a), or 
A(b) $\cap$ B(a) - see Bird & Hadzi-Pavlovic (2005). Table 2).

```{r, between_subjects_contrasts}

btwn_cont <- list(
  "Sleep_Dep" = c(0.5, -0.5, 0.5, -0.5),
  "NVH" = c(0.5, 0.5, -0.5, -0.5)
)
```

Now get an emmens object and apply the contrasts

```{r contrasts}

emms <- emmeans(mod, c("sleep_deprivation", "nvh"))
mfx <- contrast(emms, btwn_cont)
```
We also want the interaction contrast. Note there was nothing stopping us
adding it to the list of contrasts above.

```{r interaction_contrast}

int <- contrast(emms, list("int"=c(0.5, -0.5, -0.5, 0.5)))

```

We also want the simple effects of NVH at each level of sleep deprivation, and vice versa:

```{r nvh_at_sleep_dep}


nvh_by_sleep_dep <- emmeans(mod, ~ nvh|sleep_deprivation)
simp_fx_contrast  <- list(
  "NVH" = c(1, -1)
)
nvh_simp_fx <- contrast(nvh_by_sleep_dep, method=simp_fx_contrast)
nvh_simp_fx
```


```{r nvh_at_sleep_dep}


sleep_dep_by_nvh <- emmeans(mod, ~ sleep_deprivation|nvh)

sdep_simp_fx <- contrast(sleep_dep_by_nvh, method=simp_fx_contrast)
sdep_simp_fx

```

## Simultaneous confidence intervals

Now we apply the SMR correction to all our contrasts, and get the confidence intervals for all of them together.

```{r}


all_fx <- c(mfx, int, nvh_simp_fx, sdep_simp_fx) 
family_list = as.list(rep("b", length(all_fx)))
alphas = 0.05 # because we adjust by contrast

# because we are testing for A(b) and B(a) (and implicitly for A, B, and AB)
# we define our smr parameters using row 4 of Table 2 of Bird & Hadzi-Pavlovic (2005)

p <- 2 
q <- 2
smr_params <- list(
  p = p,
  q = q,
  n_sim = 100000, # this is also the default
  seed = 42 # set seed if you want to replicate simulation of smr distribution
)

psyci(model=mod, 
      contrast_tables = all_fx, 
      method = "smr", 
      smr_params = smr_params,
      family_list = family_list, 
      between_factors = list("sleep_deprivation", "nvh"),       
      alpha = alphas)

```


Info from Bird & Hadzi-Pavlovic (2005). 
See page 359 for SMR parameters.
See p 360 for the conversion of SMR into a critical constrant.




