---
title: "between_subjects_2x2_factorial-w-simple"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{between_subjects_2x2_factorial-w-simple}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
library(PsyR)
library(tidyverse)
library(emmeans)
library(afex)
library(forcats)
```

Following the example in chapter 4 of Bird (2004) Analysis of Variance via Confidence Intervals, of a coherent analysis for a 2 x 2 between subjects design with simple effects.

## Load the data

Lets load the data, make it longform, and define the factors 

```{r loaddata}

data(drive_sleepy)
head(drive_sleepy, 5)
```

```{r longform}

drive_sleepy <- drive_sleepy %>% 
  mutate(sleep_deprivation = factor(sleep_deprivation),
         sleep_deprivation = fct_recode(sleep_deprivation, "None" = "0", "High" = "12"),
         sleep_deprivation = fct_relevel(sleep_deprivation, "High", "None"),
         nvh = factor(nvh),
         nvh = fct_relevel(nvh, "high", "low"))

```

## Model 

Now run the anova

```{r model}

# perform the statistical model
mod <- aov_ez("subID", 
              "performance", 
              drive_sleepy, 
              between = c("sleep_deprivation", "nvh"))

```

## Contrasts

Lets define and obtain our contrast tables

```{r, between_subjects_contrasts}

btwn_cont <- list(
  "Sleep_Dep" = c(0.5, -0.5, 0.5, -0.5),
  "NVH" = c(0.5, 0.5, -0.5, -0.5)
)

```

Now get an emmens object and apply the contrasts

```{r contrasts}

emms <- emmeans(mod, c("sleep_deprivation", "nvh"))
mfx <- contrast(emms, btwn_cont)
```
We also want the interaction contrast:

```{r interaction_contrast}

int <- contrast(emms, list("int"=c(0.5, -0.5, -0.5, 0.5)))

```

We also want the simple effects of NVH at each level of sleep deprivation, and vice versa:

```{r nvh_at_sleep_dep}


nvh_by_sleep_dep <- emmeans(mod, ~ nvh|sleep_deprivation)
simp_fx_contrast  <- list(
  "NVH" = c(1, -1)
)
nvh_simp_fx <- contrast(nvh_by_sleep_dep, method=simp_fx_contrast)
nvh_simp_fx
```


```{r nvh_at_sleep_dep}


sleep_dep_by_nvh <- emmeans(mod, ~ sleep_deprivation|nvh)

sdep_simp_fx <- contrast(sleep_dep_by_nvh, method=simp_fx_contrast)
sdep_simp_fx

```

## Simultaneous confidence intervals

Now we apply the SMR correction to all our contrasts, and get the confidence intervals for all of them together.

```{r}


all_fx <- c(mfx, int, nvh_simp_fx, sdep_simp_fx) 
family_list = list(rep("b", times=length(all_fx))) 
alphas = 3*.05 # 3 * alpha / nk

# because we are using the SMR distribution for product
# contrasts, we have to specify come extra parameters 
# on page 359 in Table 2
ps <- c(NA, 1, 1, 1) # the latter 2 are potentially 2
qs <- c(NA, 1, 2, 2)
nu1 <-c(NA, 1, 2, 2) # the latter 2 are potentially 3
smr_params <- list(
  p = ps,
  q = qs,
  nu1 = nu1
)

psyci(model=mod, 
      contrast_tables = all_fx, 
      method = "smr", 
      smr_params = smr_params,
      family_list = family_list, 
      between_factors = list("sleep_deprivation", "nvh"),       
      alpha = alphas,
      independent = FALSE)

```


Info from Bird & Hadzi-Pavlovic (2005). 
See page 359 for SMR parameters.
See p 360 for the conversion of SMR into a critical constrant.




