---
title: "F-STP-and-Scheffe-CI-posthoc"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{F-STP-and-Scheffe-CI-posthoc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Title

Info

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Loading packages

To use Psy to calculate confidence intervals over contrast effects, we will use a workflow that uses functions from [afex](https://www.rdocumentation.org/packages/afex/versions/1.4-1) to implement ANOVA models, and [emmeans](https://rvlenth.github.io/emmeans/) to obtain estimated marginal means and the required contrast estimates. So let's load those packages, and any others that are helpful for our workflow.

```{r, loadpackages}

library(afex)
library(emmeans)
library(tidyverse)
library(ggplot2)
library(PsyR)

```

## The data

Lets load the data, then call the `head()` function to print the first 5 rows.

```{r loaddata}

data(fatigue)
head(fatigue, 5)
```

Have you forgotten the details of the fatigue dataset? Call the help function using `?data` to open its documentation.

```{r, whatfatigue}

?fatigue
```

Time for some tidying! `mutate()` creates a new column called "Group", then assigns the labels of "Alert", "Mild", "Moderate" and "Extreme" to the previously coded group levels "1", "2", "3", "4". We have also called the `factor()` function, assigning our new column as a factor rather than a double. This will be helpful later when we visualise our data, not to mention that it makes for some more human readable category names!
```{r sortfatigue}

fatigue <- fatigue %>%
             mutate(Group = factor(Group, 
                                  levels=c(1,2,3,4),
                                  labels=c('Alert', 
                                           'Mild', 
                                          'Moderate',
                                          'Extreme')))
```

Time to plot our data, with help from the ggplot2 package.  
First, using a pipe, lets tell `ggplot()` to use the fatigue dataset (the one we just edited, with the better labels). Then, within `aes()` lets create a canvas with "Group" on the X axis and "Errors" on the Y axis. Lets also fill our boxplots based on the Group they represent. 

Next, lets create a boxplot `geom_boxplot()` and make it look nice and clean using `theme_classic()`.
#' plot not wide enough
```{r boxplotit}

fatigue %>% ggplot(aes(x = Group, y = Errors, fill = Group)) +
  geom_boxplot() + theme_classic()

```

## The Analysis

Lets run some statistical tests! First, lets specify that we want a multivariate analysis, using the `afex_options()` function. Next, lets create a new object *mod*, taking in the ID of our subject, their errors score and their fatigue level. Next, lets specify that our data is between-subjects, as identified by our "Group" column. Finally, lets create a new object *emms* (short for estimated marginal means) calling the `emmeans()` function and specifying that we want to take in *mod* and output based upon "Group".

```{r, aov}

# set emmeans option to multivariate
afex_options(emmeans_model = "multivariate")
# perform the statistical model
mod <- aov_ez("id", 
              "Errors", 
              fatigue, 
              between = "Group")

emms <- emmeans(mod, "Group")
emms

```

Time to define our contrasts! Lets create an object called *between_contrasts*. Call the `list` function, then create four contrasts. The contrasts represent the following:  
* "Fatigue-Alert" - The difference between the average of fatigued conditions and the alert condition  
* "Extreme-Mod,Mild" - The difference between Extreme fatigue and the average of Moderate and Mild fatigue.  
* "Low-Hi" - The difference between the average of Alert and Mild fatigue versus the average of Moderate and Extreme fatigue.  
* "Alert-Mild" - The difference between the Alert and Mild fatigue condtions.

```{r definecontrasts}
### note to selves - is there a way to get emmeans to do the scaling here?

between_contrasts = list(
  "Fatigue-Alert" = c(-1, 1/3, 1/3, 1/3),
  "Extreme-Mod,Mild" = c(0, -1/2, -1/2, 1),
  "Low-Hi" = c(1/2, 1/2, -1/2, -1/2),
  "Alert-Mild" = c(1, -1, 0, 0)
)
between_contrasts
```

Lets create an ANOVA table. Create the *btwn_fx* object, then call the `contrast` function, setting *emms* in the "object" arguement and *between_contrasts* in the "future_sadie" arguement.

```{r, between_fx}

btwn_fx <- contrast(emms, between_contrasts)
btwn_fx
```

Finally, lets (future_sadie_here)

```{r, psyci}


contrasts_w_cis = psyci(mod, contrast_tables=list(btwn_fx),  family_list=list("b"), method="ph", alpha=.05, between_factors = list("Group"))

```
