---
title: "GCR-probability-calculator"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GCR-probability-calculator}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(PsyR)
```

## Roy's Largest Root

The first step is to get Roy's Largest Root. To do this we need to fit our model as a MANOVA

Let's use the built in `social_anxiety` dataset to run our analysis


```{r load}

data(social_anxiety)
head(social_anxiety, 5)
```

Pivot the data to wideform

```{r, wideform}

social_anxiety  <- social_anxiety %>% 
  pivot_wider(names_from = Session, values_from = Score)

```

Now we run a MANOVA analysis to get Roy's GCR of the co-variance matrix

```{r, multimod}

fit <- manova(cbind(Pre, Post, FU1, FU2) ~ Group, data = social_anxiety)
summary(fit, test = "Roy")

```

So we are working with a value of 0.3903

## compute probability

Now we compute the probability of observing a GCR of 0.39 of our analysis. 

```{r, set params}

theta = 0.3903
v_w = 3 # df within = number of levels of the within subject factor - 1
v_b = 2 # df between = number of levels of the between subject factor - 1
v_e = 87 # df error = total number of subjects - number of groups


```

Now the parameters are set, we can observe the probability of theta given v_w, v_b, and v_e

```{r, compute prob}

p = pgcr(theta, v_w, v_b, v_e)
p
```
