---
title: "J-K-Design_inference-on-simple-effects"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{J-K-Design_inference-on-simple-effects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Title

Info

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Loading packages

To use Psy to calculate confidence intervals over contrast effects, we will use a workflow that uses functions from [afex](https://www.rdocumentation.org/packages/afex/versions/1.4-1) to implement ANOVA models, and [emmeans](https://rvlenth.github.io/emmeans/) to obtain estimated marginal means and the required contrast estimates. So let's load those packages, and any others that are helpful for our workflow.

```{r, loadpackages}

library(afex)
library(emmeans)
library(tidyverse)
library(ggplot2)
library(PsyR)
```

## The data

```{r loaddata}

data(experience)
head(experience, 5)
```

Sort out the factors

```{r sortfatigue}

experience <- experience %>%
                mutate(Experience = factor(Experience,
                          levels=c("10yrs", "3yrs", "<1yr")),
                       Therapy = factor(Therapy,
                                        levels=c("Behavioural",
                                                 "Cognitive",
                                                 "Psychotherapy",
                                                 "Group")))
```


```{r plotdata}

experience %>% ggplot(aes(x=Therapy, y=Outcome, fill=Experience)) +
  geom_boxplot() + theme_minimal()

```

```{r, aov}

# perform the statistical model
mod <- aov_ez(id="SubjID", 
              dv="Outcome", 
              data=experience, 
              between = c("Experience", "Therapy"))

emms <- emmeans(mod, c("Experience", "Therapy"))
emms

```

First, define the factor A (Experience of therapist) contrasts

```{r btwn_contrasts}

experience_contrasts = list(
  "lots_vs_little" = c(0, 1, -1),
  "lots_vs_some" = c(1, -1, 0)
)

therapy_contrasts = list(
  "evidence_vs_woowoo" = c(0.5, 0.5, -0.5, -0.5),
  "behav_vs_cog" = c(1, -1, 0, 0),
  "woo1_vs_woo2" = c(0, 0, 1, -1)
)  

```

## First, let's sort the A and A(b) and AB family

Calculate the main effects for factor A

```{r, factor_a_fx}

emms_exp <- emmeans(mod, c("Experience"))
exp_fx <- contrast(emms_exp, experience_contrasts)
exp_fx

```

Now let's get the A(b) effects

```{r, A_b}

Exp_g_therapy <- emmeans(mod, ~ Experience|Therapy)
Ab_fx <- contrast(Exp_g_therapy, method=experience_contrasts)

```

And last, the AB effects

```{r, AB}

interaction_fx = contrast(emms, interaction=list(experience_contrasts,
                                                 therapy_contrasts))
interaction_fx
```

For a priori, bonferroni method:

Now we have our 3 sets of contrasts (all appropriately scaled).
The Bonferroni adjustment must be made at 
alpha = 0.5x2 for the A(B) family (or alpha=0.05x3, if using the all factorial model). 
alpha is divided by k, which = number of planned contrasts in the family.
Note that the A(B) family includes A simple effects contrasts, A main effects contrasts, and AB interaction contrasts.
v2 = N-JK (df error)
we can feed them into psyci to get our preferred confidence intervals.

ME (A) exp: k = 2 alpha = .1
A(b) exp @ each level of therapy k = 8, alpha = .1
AB interaction k = 6, alpha = .1
Thus k = 16, alpha = .1, for all families (as they are actually one family, as all non-independent)


```{r, get_cis}

all_fx <- c(exp_fx, Ab_fx, interaction_fx)
family_list = list("b") # to be documented
alphas = rep(list(2*.05), times=3)

psyci(model=mod, contrast_tables = all_fx, method = "bf", family_list = list("b"), 
      between_factors = list("Therapy", "Experience"), alpha = alphas,
      independent = FALSE)

```




