---
title: "manual-calc_crit-values"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{manual-calc_crit-values}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(PsyR)
```

Sometimes you just want to manually calculate a critical value yourself, and the 
consequent confidence intervals. Here is how you do that:

We are going to assume a contrast SE of 2.5, and an alpha rate of 0.05.
Let's also assume a 2 (within) x 2 (between) design, with 25 participants in each group.
Let's also just assume:
df within of 1,
a df between of 2, 
and a df error of 98.

```{r, setup vars}

v_w <- 1
v_b <- 2
v_e <- 98
SE = 2.5
```

First lets calculate the critical constant for each method:

Individual t:

```{r ind t}

cc_ind_t(v_e)

```

Bonferroni corrected t, assuming 2 contrasts in the family of whichever contrasts...

```{r bonf t}

k = 2
cc_bonf_t(v_e=v_e, n_k = 2)

```

Post-hoc Scheffe adjustment for the between family

```{r, ph_b}

cc_ph_b(v_b=v_b, v_e=v_e)

```


Or post-hoc Scheffe adjustment for the within family

```{r, ph_w}

cc_ph_w(v_w=v_w, v_e=v_e)
```

or Between x within post-hoc using Roy's GCR

```{r, ph_bw}

cc_ph_bw(v_w, v_b, v_e)
```

Let's assume we'll use this last CC to construct our confidence intervals. We can do that as follows:

```{r, cis}

cc <- cc_ph_bw(v_w, v_b, v_e)
compute_contrast_ci(cc, SE)
```


Thus the CIs would be the mean, plus and minus the above number.
