---
title: "planned-contrasts_between-within"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{planned-contrasts_between-within}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(PsyR)
```

## Load the data

```{r loaddata}

data(depression)
head(depression, 5)
```


and take a look at it

```{r sortdep}

depression <- depression %>%
               mutate(Group = fct_relevel(Group, 'Ctrl', 'Tmt1', 'Tmt2'),
                      Time = fct_relevel(Time, 'Pre', 'Post', 'FU'),
                      Subject = as_factor(Subject))
```

Now that we have a better idea about what the data is, let's make a plot to take a look at the impact of treatment on happiness scores over time.

```{r plothappiness}

depression %>% group_by(Group, Time) %>%
  summarise(mean_happiness = mean(Happiness)) %>%
  ggplot(aes(x=Time, y=mean_happiness, 
             group=Group, colour=Group)) +
  geom_line() + theme_classic()
```


Run b*w anova model

```{r aov}

# set emmeans option to multivariate
afex_options(emmeans_model = "multivariate")
# perform the statistical model
mod <- aov_ez("Subject", 
              "Happiness", 
              depression, 
              within = "Time", 
              between = "Group")

emms <- emmeans(mod, c("Group", "Time"))
emms
```


## Now define contrasts of interest

### Between contrasts

```{r, btwn}

between_contrasts = list(
  "Ts_v_Ctrl" = c(-1, 0.5, 0.5),
  "T1_v_T2" = c(0, 1, -1)
)
between_contrasts

```

Now apply the between contrasts to our model

```{r, getbtwn}

emms_grp <- emmeans(mod, c("Group"))
btwn_fx <- contrast(emms_grp, between_contrasts)
btwn_fx

```

### Within contrasts

Now do the same for within -

```{r, btwn}

within_contrasts = list(
  "Post_vs_Pre" = c(-1, 1, 0),
  "FU_vs_Post" = c(0, -1, 1)
)
within_contrasts

```

and get the within fx

```{r wthn}

emms_win <- emmeans(mod, "Time")
win_fx <- contrast(emms_win, within_contrasts)
win_fx

```

### Interaction contrasts

```{r int_fx}

emms_int <- emmeans(mod, c("Group", "Time"))
int_fx <- contrast(emms_int, interaction=list(between_contrasts,
                                              within_contrasts))
int_fx

```

## Confidence!

Now we're ready to use PsyR to get our simultaneous CIs -

```{r psyci}

psyci(model=mod, contrast_tables = c(btwn_fx, win_fx, int_fx), 
      method = "bf", family_list = list("b", "w", "bw"), 
      within_factors = list("Time"),
      between_factors = list("Group"))

```
