---
title: "post-hoc_within-subjects"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{post-hoc_within-subjects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE, warning = FALSE}
library(PsyR)
library(tidyverse)
library(emmeans)
library(afex)
```

Following the example in chapter 6 of [@bird2004] Analysis of Variance via Confidence Intervals

## Load the data

Lets load the data, make it longform, and define the factors

```{r loaddata}
data(kevin_within)
head(kevin_within, 5) %>%
  knitr::kable(format = "html", table.attr = "style='width:50%;'")
```

```{r longform}
kevin_within <- kevin_within %>%
  pivot_longer(c("Pre", "Post", "FU"),
    values_to = "Score",
    names_to = "Session"
  )

kevin_within <- kevin_within %>%
  mutate(
    Session = fct_relevel(
      Session, 
      "Pre", 
      "Post", 
      "FU"
    )
  )
```

## Model

Now run the anova

```{r model}
# set emmeans option to multivariate
afex_options(emmeans_model = "multivariate")

# perform the statistical model
mod <- aov_ez(
  "Subject",
  "Score",
  kevin_within,
  within = "Session"
)

emms <- emmeans(mod, "Session")
knitr::kable(emms, digits = 3)
```

## Contrasts

Now lets define and compute contrasts of interest

```{r phoc}
contrasts <- list(
  "Post-Pre" = c(-1, 1, 0),
  "FU-Post"  = c(0, -1, 1),
  "FU-Pre"   = c(-1, 0, 1)
)

cnt_res <- contrast(emms, contrasts)
knitr::kable(cnt_res, digits = 3)
```

## Simultaneous Confidence Intervals

Now lets compute post-hoc 95% CIs for a within subjects design

```{r, scis}
psyci(
  model = mod, 
  contrast_tables = cnt_res, 
  method = "ph", 
  family_list = list("w"),
  within_factors = list("Session")
) %>%
knitr::kable(digits = 3)
```

## The Interpretation

to be added

------------------------------------------------------------------------

logo here

------------------------------------------------------------------------
