---
title: "post-hoc_mixed-design"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{post-hoc_mixed-design}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(PsyR)
library(tidyverse)
library(emmeans)
library(afex)
```

Following the example in chapter 7 of Bird (2004) Analysis of Variance via Confidence Intervals

## Load the data

Lets load the data, make it longform, and define the factors 

```{r loaddata}

data(social_anxiety)
head(social_anxiety, 5)
```

```{r longform}

social_anxiety <- social_anxiety %>% 
  mutate(Session = fct_relevel(Session, 'Pre','Post','FU1','FU2'),
         Group = factor(Group),
         Group = fct_recode(
           Group,
           "New" = "1",
           "Standard" = "2",
           "MinContact" = "3"
         ))

```

## Model 

Now run the anova

```{r model}

# set emmeans option to multivariate
afex_options(emmeans_model = "multivariate")
# perform the statistical model
mod <- aov_ez("Subject", 
              "Score", 
              social_anxiety, 
              within = "Session",
              between = "Group")

```

## Contrasts

Lets define and obtain our contrast tables

```{r, win}

win_contrasts <- list(
  "Pre-rest" = c(1, -1/3, -1/3, -1/3),
  "Post-FUs" = c(0, 1, -0.5, -0.5),
  "FU1-FU2" = c(0, 0, 1, -1)
) 

emms_win <- emmeans(mod, "Session")

win_fx <- contrast(emms_win, win_contrasts)
win_fx

```

```{r, btwn}

btwn_contrasts <- list(
  "Ts-C" = c(0.5, 0.5, -1),
  "NT-ST" = c(1, -1, 0)
) 

emms_btwn <- emmeans(mod, "Group")

btwn_fx <- contrast(emms_btwn, btwn_contrasts)
btwn_fx

```


```{r, int}


emms_int <- emmeans(mod, c("Group", "Session"))

int_fx <- contrast(emms_int, interaction=list(btwn_contrasts, win_contrasts))
int_fx
```


## A Priori Confidence

Here we replicate the bonferroni correction first presented in chapter 7

```{r, psyci}

all_fx <- c(win_fx, btwn_fx, int_fx)
family_list = list("w", "b", "bw") 
alphas = .05 # 3 * alpha / nk

psyci(model=mod, contrast_tables = all_fx, method = "bf", family_list = family_list, 
      between_factors = list("Group"), within_factors=list("Session"), alpha = alphas)

```



Next, we run the exact same but using the post-hoc method

```{r, psyci}

psyci(model=mod, contrast_tables = all_fx, method = "ph", family_list = family_list, between_factors = list("Group"), within_factors=list("Session"), alpha = alphas)

```
