---
title: "post-hoc_within-subjects"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{post-hoc_within-subjects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

##The Packages

To use Psy to calculate confidence intervals over contrast effects, we will use a workflow that uses functions from [afex](https://www.rdocumentation.org/packages/afex/versions/1.4-1) to implement ANOVA models, and [emmeans](https://rvlenth.github.io/emmeans/) to obtain estimated marginal means and the required contrast estimates. So let's load those packages, and any others that are helpful for our workflow.

```{r setup, message = FALSE, warning = FALSE}
library(PsyR)
library(emmeans)
library(afex)
library(tidyverse)
```

Insert new vignette description here.

Following the example in chapter 6 of @bird2004 Analysis of Variance via Confidence Intervals

## The Data

Lets load the data, then call the `head()` function to print the first 5 rows.

```{r loaddata}
data(kevin_within)
head(kevin_within, 5) %>%
  knitr::kable(format = "html", table.attr = "style='width:50%;'")
```

Have you forgotten the details of the dataframe? Call the help function using `?df` to open its documentation.

```{r, whatfatigue, message = FALSE, warning = FALSE}
?kevin_within
```

Lets tidy our data! First, call `pivot_longer()`, using `values_to` to move the scores from the "Pre", "Post" and "FU" (Follow Up) conditions into a single column labelled "Score". Lets also create a column to identify what session the scores come from, using `names_to` to populate the column with our predefined session names.

Next, lets re-organise our new "Session" column. We can order the scores according to the chronological order of our data, such that all "Pre" scores are displayed at the top of the dataframe, followed by "Post" and then "FU".

```{r longform}
kevin_within <- kevin_within %>%
  pivot_longer(c("Pre", "Post", "FU"),
    values_to = "Score",
    names_to = "Session"
  )

kevin_within <- kevin_within %>%
  mutate(
    Session = fct_relevel(
      Session, 
      "Pre", 
      "Post", 
      "FU"
    )
  )
```

Time to plot our data, with help from the ggplot2 package.

Using a pipe, lets tell `ggplot()` to take in the kevin_within dataset (the one we just tidied up into longform). Then, within `aes()` lets create a canvas with "Session" on the X axis and "Score" on the Y axis. Lets also colour-code our four groups by calling fill: [Pre]{style="color: OrangeRed;"}, [Post]{style="color: OliveDrab;"}, [FU]{style="color: Blue;"}.

```{r boxplotit, fig.width = 7.1, fig.height = 4}
kevin_within %>%
  ggplot(aes(x = Session, y = Score, fill = Session)) +
  geom_boxplot() +
  theme_classic()
```

## The Analysis

Lets prepare to run some statistical tests!

Here we will create the building blocks of our eventual ANOVA output. First, lets specify that we want a multivariate analysis, using the `afex_options()` function. Next, lets create a new object *mod*. To create *mod*, call the `aov_ez()` function which takes in the "Subject" column to identify subjects, the "Score" column as the dependent variable, the kevin_within dataframe, and the "Group" column set as between-subjects. This aggregates across our Session column, so it isn't useful on its own.

Next, lets create a new object *emms* (short for estimated marginal means) calling the `emmeans()` function and specifying that we want to take in *mod* (the object we just made) and output based upon "Session". Now we are heading in the right direction, as our results are split by factor ("Pre", "Post", "FU").
  

```{r model}
# set emmeans option to multivariate
afex_options(emmeans_model = "multivariate")

# perform the statistical model
mod <- aov_ez(
  "Subject",
  "Score",
  kevin_within,
  within = "Session"
)

emms <- emmeans(mod, "Session")
knitr::kable(emms, digits = 3)
```

Now lets define and compute our contrasts of interest. Specifically, the contrasts compare the following:

* "Post-Pre" - The difference between scores taken prior to intervention and after the intervention.

* "FU-Post" - The difference between scores taken post intervention and at a 6 month follow up.

* "FU-Pre" - The difference between scores taken prior to intervention and at a 6 month follow up.

Next, lets create our ANOVA table. Lets make the *cnt_res* object, then call the `contrast()` function from the emmeans package, setting *emms* in the "object" argument and our newly defined *contrasts* in the "method" argument.

```{r phoc}
contrasts <- list(
  "Post-Pre" = c(-1, 1, 0),
  "FU-Post"  = c(0, -1, 1),
  "FU-Pre"   = c(-1, 0, 1)
)

cnt_res <- contrast(emms, contrasts)
knitr::kable(cnt_res, digits = 3)
```

Finally, lets compute our post-hoc 95% CIs for our within-subjects design. To do so, lets call `psyci()` from the PsyR package, specifying the following:

* The model is specified as the previously made *mod* object, where the model must be made using the afex package - afex_aov

* The contrast_tables argument is set as the previously made *cnt_res*  

* The method is set as "ph", creating confidence intervals using a post-hoc method (estimate +- SE x cc)  

* The family list is set as "w", specifying that the contrast table is within-subjects only

* The within_factors argument is set according to our desired group names, as in the "Group" column
  

```{r, scis}
psyci(
  model = mod, 
  contrast_tables = cnt_res, 
  method = "ph", 
  family_list = list("w"),
  within_factors = list("Session")
) %>%
knitr::kable(digits = 3)
```

## The Interpretation

There is insufficient evidence to suggests a significant difference exists in any of our contrasts of interest (p > 0.05). However, our SCIs can give us some insight into the directionality of our data.

* CHECK p-val being less than 0.05 but still lower and upper inconsistent?? Might have to cast mind back to this in 3001 this may be an ignore p vals moment.

* Compared to post-intervention, scores at 6 month follow up show large variance. Participants at follow up compared to post may have improved by up to 16.636 units, or may have regressed by 10.636 units.

* Compared to pre-intervention, scores at 6 month follow up show large variance. Participants at follow up compared to pre may have improvde by up to 17.836 units, or may have regressed by 5.836 units.

is something fishy going on here?

------------------------------------------------------------------------

logo here

------------------------------------------------------------------------
