---
title: "psyci-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{psyci-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo = FALSE, results = "hide", message = FALSE}
library("tidyverse")
library("afex")
library("PsyR")
library("emmeans")
knitr::opts_chunk$set(fig.width = 4.5, class.output = "ro")
```


<!-- @index Vignettes!psyci -->

## Contents {#contents}

  1. [Installation](#installation) 
  2. [Example](#example) 


## Installation {#installation}

You can install the development version of PsyR from [GitHub](https://github.com/) with:


```{r, results = "hide"}
# install.packages("devtools")
# devtools::install_github("garner-code/PsyR")
```
## Example

This is a basic example which shows you how to obtain between subject confidence 
intervals. All of the following examples will use the post-hoc method 
(see Bird, 2002, for details on the post-hoc method):

As an illustration, consider the `spacing` dataset provided with the package. The data contains ...

```{r, echo = FALSE, message = FALSE}
data(spacing)
spacing %>% 
  ggplot2::ggplot(aes(x = as.factor(spacing), y = yield, color = as.factor(group))) +
  geom_point() +
  stat_summary(fun.data = "mean_cl_boot", geom = "line", aes(group = group)) +
  coord_cartesian() + # coord
  labs(
    x = "Spacing",
    y = "Yield",
    color ="Group"
  ) 
dat <- spacing
```

First, let’s set emmeans() option to multivariate and fit a mixed ANOVA model that includes a between-subject factor `group` and a within-subject factor `spacing`:

```{r}
afex_options(emmeans_model = "multivariate")

mod <- aov_ez("subj", "yield", dat, within = "spacing", between = "group")
```

To obtaining between-subject confidence intervals, we first define between-group contrasts. 

```{r}
emm_btwn <- emmeans(mod, "group")
con_b <- list(
  "12vs34" = c(0.5, 0.5, -0.5, -0.5), # groups 1 & 2 vs groups 3 & 4
  "1vs2" = c(1, -1, 0, 0), # and so on...
  "3vs4" = c(0, 0, 1, -1)
)
btwn_con <- contrast(emm_btwn, con_b)
```

After setting the contrast tabel `con_b`, feed it into psyci() with the chosen method, family, and factor, names. Methods can be found in ?psci().
```{r}
psyci(model=mod, contrast_table = btwn_con, method="ph", family="b", between_factors = list("group"))
```

Within-subject confidence intervals can also be defined as follows:
```{r}
emm_win <- emmeans(mod, "spacing")
con_w <- list(
  "20vs40" = c(1, -1, 0),
  "20vs60" = c(1, 0, -1),
  "Quad" = c(0.5, -1, 0.5)
)
con_win <- contrast(emm_win, con_w)
```

The also feed it into psyci() to generate 95% CIs for the within subjects contrasts.
```{r}
psyci(model=mod, contrast_table = con_win, method="ph", family="w", 
      within_factors = list("spacing"))
```

And finally, here is an example of how to generate CIs for between x within 
contrasts, using the same post-hoc method.
```{r}
# get emms for each cell from the between x within design
emm_int <- emmeans(mod, c("group", "spacing"))

# sometimes defining the between x within contrast is a little more involved,
# but here is one way. You need to wind up with a contrast vector that has
# as many elements as there are rows in your table of emms for each cell in 
# the between x within design
n_within = 3
n_group = 4
win_full <- lapply(con_w, function(x) rep(x, each = n_group))
btwn_full <- lapply(con_b, function(x) rep(x, times = n_within))
con_i <- lapply(btwn_full, function(x) lapply(win_full, function(y) x * y))
con_int <- contrast(emm_int, con_i)

# check mean difference option in Psy (go back to win app and doc)
# potentially add a scheffe function 
# add p values from gcr etc procedures to the table
# generate 95% CIs for the between x within subjects contrasts
psyci(model=mod, contrast_table = con_win, method="ph", family="bw", 
      within_factors = list("spacing"), between_factors=list("group"))

```


Another example with a dataset of three-level repeated measures. 
```{r, echo = FALSE, message = FALSE}
data(priming)

priming %>% 
  ggplot2::ggplot(aes(x = as.factor(Prime), y = ReactionTime)) +
  geom_point() +
  geom_smooth(method = lm, se = TRUE) +
  coord_cartesian() + # coord
  labs(
    x = "Prime",
    y = "Reaction Time"
  ) 
dat_3level <- priming
```

```{r}
afex_options(emmeans_model = "multivariate")

mod_3level <- aov_ez("Subject", "ReactionTime", dat_3level, within = "Prime")
emm_within <- emmeans(mod_3level, "Prime")
withinContrasts = list(
"Other-Stereo" = c(-2, 1, 1)/2,
"Atyp-Neut" =c(0, 1, -1)
)
con_within <- contrast(emm_within, withinContrasts)
psyci(model = mod_3level, contrast_table = con_within, method="ph", family="w", within_factors = list("Prime"))
```

